GeckoGraph: A Visual Language for Polymorphic Types

### 2025-04-09 main.tex
Image float
Game levels

\begin{figure}[]
  \includegraphics[width=\linewidth]{figures/Indicator}
  \caption{
        \label{fig:indicator}
        An example of visual hierarchy of the function indicator. The function indicator can be used to easily identify the arity of a function type by counting the connecting function indicators. For higher-order functions where functions are arguments or return values of other functions, it is very easy to see the ``level" of functions. 
  }
\end{figure}

@@

\subsection{An example level}
We illustrate the task of the user study using level 4 of the game. At this level (Fig. \ref{fig:level-example}), the programmers are tasked with implementing the function \texttt{ zeroToHero :: Zero a b -> Hero b b}. The available functions are \texttt{f1::Zero a b -> Hero b a}, \texttt{f2::Zero a a -> Hero a a}, \texttt{f3::Zero a b -> Hero b a}, and \texttt{f4::Zero a b -> Hero b b}. Two generic functions \texttt{(\$)} and \texttt{(.)} are provided to improve the ergonomics of composing functions, but all tasks can be completed without the use of generic functions. The possible solution and other details of the level can be found in the appendix (Appendix \ref{levels}).

\begin{figure}[h]
  \includegraphics[width=\linewidth]{figures/GamePlay}
  \caption{\label{fig:level-example} A screenshot from  the game ZeroToHero. On this level -- level 4 (Shown at F) -- the players need to implement the function \texttt{ zeroToHero :: Zero a b -> Hero b b} (B). They write their own definitions in the code editor (D) using a set of provided functions (G). The inferred type of their current definition is shown in (C). When ready, they can test their solution by clicking on the \textit{Attempt} button (A). They can also skip a level by clicking on the \textit{Bypass} button next to it. The output from the compiler, if there is any, is shown in a window below (E). The GeckoGraph in the screenshot uses a different color scheme that is optimized for computer screens.}
\end{figure}

At each level of the game, the programmer must select the right set of functions and compose them correctly to achieve the target result. Not all provided functions are useful. In particular, for this level, participants must discover that only \texttt{f4} and \texttt{f2} are necessary to produce the desired results. One implementation that satisfies the target type is \texttt{zeroTohero z = f2 (f4  z)}.

@@

Levels 5, 7, and 8 are the only three levels that include functions from standard Haskell library, baring the \texttt{(.)} and \texttt{(\$)} provided for convenience. Level 3 requires programmers to use the \texttt{fst} and \texttt{snd} functions to extract value from a tuple. Level 7 requires programmers to use the \texttt{(<*>)} function of the \texttt{Applicative} class, while level 8 requires the \texttt{fmap} function of the \texttt{Functor} class. The authors speculate that the more experienced participants are much more familiar with these functions, hence the strong contrast on these three levels.

>>>>

\begin{figure}[h]
  \includegraphics[width=\linewidth]{figures/Indicator}
  \caption{
        \label{fig:indicator}
        An example of visual hierarchy of the function indicator. The function indicator can be used to easily identify the arity of a function type by counting the connecting function indicators. For higher-order functions where functions are arguments or return values of other functions, it is very easy to see the ``level" of functions. 
  }
\end{figure}

@@

\subsection{Example Levels}
\begin{figure}[h]
  \includegraphics[width=\linewidth]{figures/Level2}
  \caption{\label{fig:level-example} A screenshot from  the game ZeroToHero. On this level -- level 2 (Shown at A) -- the players need to implement the function \texttt{ zeroToHero :: Zero a -> Hero a} (B). They write their own definitions in the code editor (D) using a set of provided functions (G). The inferred type of their current definition is shown in (C). When ready, they can test their solution by clicking on the \textit{Attempt} button (E). They can also skip a level by clicking on the \textit{Skip (n left)} button next to it (F).}

\end{figure}

We illustrate the task of the user study using level 2 of the game. At this level (Fig. \ref{fig:level-example}), the programmers are tasked with implementing the function \texttt{ zeroToHero :: Zero a -> Hero b}. The available functions are \texttt{runZero::Zero a -> a}, \texttt{makeHero:: a -> Hero a}. A generic function \texttt{(\$)} is provided to improve the ergonomics of composing functions, but all tasks can be completed without the use of generic functions. The possible solution and other details of the level can be found in the appendix (Appendix \ref{levels}).
To complete this level, player only needs to use each of the provided function once. The implementation {\tt zeroToHero z = makeHero (runZero z)} is one possible solution.

\begin{figure}[!h]
  \includegraphics[width=\linewidth]{figures/Level3}
  \caption{\label{fig:level3} A screenshot showing the level 3 of the game ZeroToHero.  In this level, the players need to implement the function \texttt{ zeroToHero :: Zero a -> Hero (a a)}. Players are helped by the functions: {\tt f1 :: Zero a -> Hero a}, {\tt f2 :: Zero a -> (a, a)}, and {\tt f3 :: Hero a -> Hero (a, a)}.}
\end{figure}

In level 3, the player need to implement the function {\tt zeroToHero :: Zero a -> Hero (a, a)}.  In this level, not all provided functions are used in the final result. In particular,  players must discover that only \texttt{f1} and \texttt{f3} are necessary to produce the desired results. One implementation that satisfies the target type is \texttt{zeroTohero z = f3 (f1 z)}. Another possible implementation is {\tt zeroToHero z = f3 . f1 \$ z}.

@@

Levels 5, 7, and 8 are the only three levels that include functions from standard Haskell library, baring the \texttt{(.)} and \texttt{(\$)} provided for convenience. Level 5 requires programmers to use the \texttt{fst} and \texttt{snd} functions to extract value from a tuple. Level 7 requires programmers to use the \texttt{(<*>)} function of the \texttt{Applicative} class, while level 8 requires the \texttt{fmap} function of the \texttt{Functor} class. The authors speculate that the more experienced participants are much more familiar with these functions, hence the strong contrast on these three levels.
